---
title: "Conditional probability using simulation"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```


#### Load the bats data to a dataframe
```{r}
dBat = read.csv('Data/bats.csv', header = TRUE, row.names = 1, stringsAsFactors = TRUE)
head(dBat)
```

##### Get the genes that are expressed for the 1st bat
```{r}
b1 = colnames(dBat)[dBat[1,1:5] > 0]
#bt1 = sum(dBat[1,1:5])
#bt1 
b1 
```

#### Probability that a random bat carries the Ebola virus
```{r}
ran=sum(dBat$Ebola == 'TRUE')
eE = ran/nrow(dBat)
eE
ebola = mean(dBat$Ebola)
ebola
```

#### Probability that gene-1 through gene-5 is expressed in a random bat
```{r}
all = apply(dBat[, 1:5], 2, mean)
all
```

#### Is expression of gene-1 through gene-5 dependent or independent of the bat carrying the Ebola virus
```{r}
# Subset the dataframe to those bats that carry the Ebola virus


# Conditional probability that gene-1 through gene-5 is expressed
# given that the bat carries the Ebola virus

```

```{r}
with = dBat[dBat$Ebola == "TRUE", ]
without = dBat[dBat$Ebola == "FALSE", ]
probwith = apply(bats_with_ebola[, 1:5], 2, mean)
probwithout = apply(bats_without_ebola[, 1:5], 2, mean)
print("Conditional probabilities for bats with Ebola:")
print(probwith)
print("Conditional probabilities for bats without Ebola:")
print(probwithout)
```

#### Probability that a random bat carries the Ebola given that gene-3 through gene-5 is expressed
```{r}
bats = dBat[dBat$Gene.3 == "TRUE" & dBat$Gene.4 == "TRUE" & dBat$Gene.5 == "TRUE", ]
ebola = sum(bats$Ebola == "TRUE")
prob = ebola / nrow(bats)
print(prob)
```

