---
title: "Intro to R"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Creating homeogenous datastructures: vectors and matrices (built on the fundamental datatypes character, numeric, integer, and logical)
```{r}
myvec=c(1,2,3)
myvec1=c('a','b','c')
myvec2=c(27.2,32.4,45,8)
myvec3=c(TRUE,FALSE,TRUE,FALSE)
mymat=matrix(c(1,2,3,4,5,6),nrow=2,ncol=3)
mymat1=matrix(c(1,2,3,4,5,6),nrow=2,ncol=3,byrow = TRUE)
print(myvec)
print(myvec1)
print(myvec2)
print(myvec3)
print(mymat)
print(mymat1)
typeof(myvec)
typeof(myvec1)
typeof(myvec2)
typeof(myvec3)
typeof(mymat)
```

#### Creating a list, a heterogenous datastructure
```{r}
mylist=list('a',23,1.00)
print(mylist)

```

#### Accessing elements of a datastructure
```{r}
myvec[1]
myvec1[3]
myvec2[2]
mymat[1,2]
```

#### Loading data into a dataframe, a heterogenous datastructure
```{r}
#file = 'http://openmv.net/file/food-texture.csv'
file = 'Data/food-texture.csv'
foodData = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
head(foodData, 5)
```

#### Attributes of a dataframe
```{r}
str(foodData)
nrow(foodData)
ncol(foodData)
colnames(foodData)
rownames(foodData)
```

#### Get the data type and the data structure associated with an object
```{r}
typeof('oil')
typeof(foodData)
```

#### Accessing elements of a data frame
```{r}
foodData$Oil
foodData[['Density']]
foodData[,'Crispy']
foodData['Fracture']
```

#### Accessing elements of a particular column of a data frame
```{r}
foodData$Oil
foodData$Density
foodData$Crispy
foodData$Fracture
foodData$Hardness
```

#### Accessing elements of multiple columns of a dataframe (using base R and dplyr)
```{r}
foodData[c('oil','Density')]
foodData %>% select(c(oil,Density))
foodData %>% select(-c(oil,Density))
```

```{r}
data='Data/court.csv'
jd=read.csv(data, header = TRUE)
head(jd,10)

```

```{r}
apply(jd, 2, mean)
```

#### Matrix example - judges problem
```{r}
courtdecision = function(n){
  jcp = c(0.9478,0.9535,0.8954,0.8977,0.4)
  jicp = 1-jcp # broadcasting
  p = matrix(c(jcp,jicp), nrow = 2, ncol = 5, byrow = TRUE)
  #p = rbind(jcp, jicp)
  decision = c(1, 0)
  cdecision = vector('integer', ncol(p))
  for (j in 1:length(cdecision)){
    cdecision[j] = sample(decision, 1, replace = FALSE, prob = p[,j])
    cdecision[5]=cdecision[1]
  }
  if (sum(cdecision) >= 3){
    return(1)
    }
  else{
    return(0)
}
}
result = courtdecision()
print(result)
```

#### Court incorrect probability
```{r}
results = replicate(10000, courtdecision())
mean(results)
print(1-mean(results))
```

#### Accessing rows of a dataframe satisfying certain conditions
```{r}
```

#### Accessing row of a dataframe satisfying certain condition and particular columns
```{r}
```

#### Accessing rows of a dataframe satisfying certain conditions
```{r}
#Filter samples with Crispy index 9 or 15 (which are very rare)
```

#### Modifying a column of a dataframe
```{r}
```

#### Modify Crispy column to reflect high (0) or low (1) crispiness
```{r}
```

##### Change Crispy and Crispylevel columns to factor (categorical) type
```{r}
categorical_cols = c('Crispy', 'Crispylevel')
foodData[categorical_cols] = lapply(?, ?)
str(foodData)
```

#### Visualize the OilPercentage feature
```{r}
```

#### In-built functions for dataframes
```{r}
# Mean oil percentage across all samples

# Mean-centering of OilPercentage

# Sum of the squared deviation from the mean

# Average of the squared deviation from the mean

# Variance of OilPercentage

# Standard deviation of OilPercentage
```

#### Scatter plot between OilPercentage and Density
```{r}
```

#### Scatter plot between OilPercentage and Density color coded with Crispylevel
```{r}
```

#### Scale continuous columns of the dataframe
```{r}
```

#### Calculate the correlation matrix for the continuous features
```{r}

```

#### Scatter plot between Density and Hardness
```{r}
```

#### Scatter plot between Density and Fracture
```{r}
```

```{r}
data1= 'Data/emergency.csv'
dfPatient = read.csv(data1, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
head(dfPatient, 5)
```

```{r}
length(dfPatient)
str(dfPatient)
nrow(dfPatient)
ncol(dfPatient)
```

```{r}
good=sum(dfPatient$Condition == 'good')
good
goodprob=good/nrow(dfPatient)
print('probability of the next incoming patient')
print('with Good Condition  is : ')
goodprob
```

```{r}
fair=sum(dfPatient$Condition == 'fair')
fair
fairprob=fair/nrow(dfPatient)
print('probability of the next incoming patient')
print('with Fair Condition  is : ')
fairprob
```

```{r}
serious=sum(dfPatient$Condition == 'serious')
serious
seriousprob=serious/nrow(dfPatient)
print('probability of the next incoming patient')
print('with Serious Condition  is : ')
seriousprob
```

```{r}
insure=sum(dfPatient$Insurance == 'insured')
insure
insuredprob=insure/nrow(dfPatient)
print('probability of the next incoming patient')
print('with Insurance  is : ')
insuredprob
```

```{r}
notinsure=sum(dfPatient$Insurance == 'not insured')
notinsure
notinsuredprob=notinsure/nrow(dfPatient)
print('probability of the next incoming patient')
print('without Insurance is : ')
notinsuredprob
```

```{r}
insure_serious=sum(dfPatient$Insurance == 'insured' & dfPatient$Condition == 'serious')
insure_serious
iserious=sum(dfPatient$Condition == 'serious')
iserious
insure_seriousprob=insure_serious/iserious
insure_seriousprob
```

```{r}
serious_insure=sum( dfPatient$Condition == 'serious'  &  dfPatient$Insurance == 'insured')
serious_insure
insure=sum(dfPatient$Insurance == 'insured')
insure
serious_insureprob=serious_insure/insure
serious_insureprob
```

```{r}
good=sum(dfPatient$Condition == 'good')
goodprob=good/nrow(dfPatient)
fair=sum(dfPatient$Condition == 'fair')
fairprob=fair/nrow(dfPatient)
serious=sum(dfPatient$Condition == 'serious')
seriousprob=serious/nrow(dfPatient)
insure=sum(dfPatient$Insurance == 'insured')
insuredprob=insure/nrow(dfPatient)
notinsure=sum(dfPatient$Insurance == 'not insured')
notinsuredprob=notinsure/nrow(dfPatient)
conditions = c("good", "fair", "serious")
insurance = c("insured","not insured")
probs = c(goodprob, fairprob, seriousprob)
inprobs=c(insuredprob,notinsuredprob)
simulated_conditions = sample(conditions, size = 10000, replace = TRUE, prob = probs)
simulated_insurance = sample(insurance, size = 10000, replace = TRUE, prob = inprobs)
simulated_patients = data.frame(Condition = simulated_conditions, Insurance = simulated_insurance)
simulated_patients
good=sum(simulated_patients$Condition == 'good')
good
notinsure=sum(simulated_patients$Insurance == 'not insured')
notinsure
insure=sum(simulated_patients$Insurance == 'insured')
insure
serious=sum(simulated_patients$Condition == 'serious')
serious
fair=sum(simulated_patients$Condition == 'fair')
fair
```


```{r}
#print('probability of the next incoming patient')
#print('with Condition Good')
#Good = dfPatient[dfPatient$Condition == 'Good', ]
#Goodcount = nrow(Good)
#GoodProb = Goodcount/nrow(dfPatient)
#GoodProb
#if(dfPatient$Insurance == 'insured'){
 # return (1)
#  }
#else{
 # return (0)
#}

#insure=sum(dfPatient$Insurance == 'insured')
#insuredprob=insure/nrow(dfPatient)
#insuredprob

```












