---
title: "Quantifying uncertainty using simulation"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Sampling space and likelihood vector
```{r}
s = c('H', 'T')
p = c(0.5, 0.5)
```


#### Function to simulate one trial of the random experiment which in this case is tossing 10 fair coins at a time
```{r}
simTrial = function(){
  return(sample(s, 10, replace = TRUE, prob = p))
}
```

#### Replicate the trials of the random experiment a large number of times
```{r}
nsimulations = 1
simulatedData = replicate(nsimulations, simTrial())
simulatedData
```

#### Function to check if event of interest has occured which in this case is getting exactly 3 heads
```{r}
checkEvent = function(data){
  return(sum(data == 'H') == 3)
}
```

#### Calculate the fraction of times the event of interest occured which is what we define as the approximate probability of the event E happening in the next trial of the random experiment
```{r}
#simulatedData
mean(apply(simulatedData, 2, checkEvent))
```

```{r}
cd = read.csv('Data/court.csv')
cd
```


```{r}
court = function(n) {
  jcp = apply(cd, 2, mean)
  jicp = 1 - jcp
  s = c(1, 0)
  decision = vector('integer', n)
  for (j in 1:n) {
    decision[j] = sample(s, 1, prob = c(jcp[j], jicp[j]))
  }
  return(decision)
}
```

```{r}
simulateData = replicate(1000, court(5))
```

```{r}
check = function(Data) {
  if(sum(Data) <=2 ){
    return(1)
  }  
  else{
    return(0)
  }
}
```

```{r}
result_check = apply(simulateData, 2, check)
mean_count = mean(result_check)
mean_count
```

```{r}
sam = c(1,2,3,4,5,6)
pro = c(1/6,1/6,1/6,1/6,1/6,1/6)
```

```{r}
c = sample(sam,1, replace = TRUE, prob = pro)
c
```

```{r}
c = sample(sam,4, replace = TRUE, prob = pro)
c
```


```{r}
sim = function(){
  c = sample(sam,4, replace = TRUE, prob = pro)
}
```

```{r}
s = replicate(1000, sim())
#s
```

```{r}
checkdie = function(Data) {
  if (sum(Data == '6') >=1) {
    return(1)
  } else {
    return(0)
  }
}
```

```{r}
mean(apply(s, 2, checkdie))
```

```{r}
ple = c(2,3,4,5,6,7,8,9,10,11,12)
lit = c(1/36, 2/36, 3/36, 4/36, 5/36, 6/36, 5/36, 4/36, 3/36, 2/36, 1/36)
```

```{r}
c = sample(ple, 1, replace = TRUE, prob = lit)
c
```

```{r}
c = sample(ple, 24, replace = TRUE, prob = lit)
c
```

```{r}
late = function(){
  c = sample(ple, 24, replace = TRUE, prob = lit)
}
```

```{r}
x = replicate(1000, late())
#x
```

```{r}
die = function(Data) {
  if (sum(Data == '12') >=1) {
    return(1)
  } else {
    return(0)
  }
}
```

```{r}
mean(apply(x, 2, die))
```

```{r}
bd = c(1:365)
c = sample(bd, 60, replace = TRUE)
c
```

```{r}
day = function(){
  c = sample(bd, 60, replace = TRUE)
}
```

```{r}
xx = replicate(1000, day())
```


```{r}
bday = function(Data) {
  if (any(duplicated(Data))) {
    return(1)
  } else {
    return(0)
  }
}
```

```{r}
mean(apply(xx, 2, bday))
```

```{r}
bd = c(1:365)
c = sample(bd, 21, replace = TRUE)
c
```

```{r}
day = function(){
  c = sample(bd, 21, replace = TRUE)
}
```

```{r}
xx = replicate(1000, day())
```


```{r}
bday = function(Data) {
  if (any(duplicated(Data))) {
    return(1)
  } else {
    return(0)
  }
}
```

```{r}
mean(apply(xx, 2, bday))
```

```{r}
d = (choose(365,60)*factorial(60)/(365^60))
d
```

```{r}
1-(choose(6,4)*factorial(4)/(6^4))
```

```{r}
1-(choose(36,)*factorial(4)/(6^4))
```

```{r}
not= (5/6)
```

```{r}

```

```{r}
pro = not^4
pro
getting = 1 - pro
getting
```

```{r}
nnot = (35/36)
npro = nnot^24
npro
gpro= 1-npro
gpro
```

