---
title: "Timeseries Project Even Semester 2025"
editor_options:
chunk_output_type: console
output:
html_document:
df_print: paged
pdf_document: default
---

```{r}
# Load essential libraries
library(ggplot2)
library(dplyr)
library(datasets)
library(forecast)
library(xts)
library(scales)
library(lubridate)
library(quantmod)
```

```{r, echo = FALSE}
# Set ggplot theme for plotting
My_Theme = theme(axis.text.x = element_text(size = 6, angle = 90, vjust = 0.5, hjust=1),
   axis.text.y = element_text(size = 9),
   axis.title.x = element_text(size = 11),
   axis.title.y = element_text(size = 11),
   plot.title = element_text(size = 12, hjust = 0.5, face = "bold"))
```

## 1. Using R's in-built function ts(), create timeseries data for a temperature sensor measuring temperature once every 5 minutes for 24 hours starting from 00:00 hours on January 1, 2025
```{r}
data = rnorm(288, 27, 8)
ts_temperature = ts(data, frequency = 288, start = c(2025, 1))
print(ts_temperature)
plot(ts_temperature)
print(time(ts_temperature))
```

## 2. Using the xts library, create an xts timeseries object using the same data above
```{r}
# Generate timestamps 
time_index = seq(from = ymd_hms("2025-01-01 00:00:00", tz = 'Asia/Kolkata'), to = ymd_hms("2025-01-01 23:55:00", tz = 'Asia/Kolkata'), by = "5 min")
# Create an xts time series object
xts_temperature = xts(data, order.by = time_index)
print(xts_temperature)
plot(xts_temperature)
print(index(xts_temperature))
```

## 3. Plot the timeseries data using ggplot
```{r}
# Convert xts object into dataframe
df_temperature = fortify(xts_temperature)
colnames(df_temperature)[2] = c('Temperature')

# Plot the time series
ggplot(data= df_temperature) +
  geom_line(aes(x = Index, y = Temperature), color = 'blue') +
  scale_x_datetime(expand = c(0,0),  labels = date_format("%H:%M", tz = 'Asia/Kolkata'), breaks = seq(min(index(xts_temperature)), max(index(xts_temperature)), by = "60 mins"))+
  labs(title = "Temperature Timeseries", x = "Time", y = "Temperature") +
  My_Theme
```

## 4. Filter temperature data from 09:00 to 14:30 and plot it
```{r}
# Define the time range for subsetting
start_time = ymd_hms("2025-01-01 09:00:00", tz = 'Asia/Kolkata')
end_time = ymd_hms("2025-01-01 14:30:00", tz = 'Asia/Kolkata')

# Subset using xts index function
xts_temperature_filtered <- xts_temperature[index(xts_temperature) >= start_time & index(xts_temperature) <= end_time]
xts_temperature_filtered
plot(xts_temperature_filtered)
```

## 5. Plot filtered temperature data from 09:00 to 14:30 using ggplot
```{r}
# Convert xts object into dataframe
df_temperature_filtered = fortify(xts_temperature_filtered)
colnames(df_temperature_filtered)[2] = c('Temperature')

# Plot the time series
ggplot(data= df_temperature_filtered) +
  geom_line(aes(x = Index, y = Temperature), color = 'blue') +
  scale_x_datetime(expand = c(0,0),  labels = date_format("%H:%M", tz = 'Asia/Kolkata'), breaks = seq(min(index(xts_temperature)), max(index(xts_temperature)), by = "30 mins"))+
  labs(title = "Filtered Temperature Timeseries", x = "Time", y = "Temperature") +
  My_Theme
```

## 5. Load the "Applicance Energy Consumption" dataset
```{r}
# Load data into a dataframe
df_energy = read.csv("Data/energydata_complete.csv")

# Convert 'date' column to POSIXct format
df_energy$date = ymd_hms(df_energy$date, tz = "UTC")

# Convert to xts object
xts_energy = xts(df_energy[, -1], order.by = df_energy$date)
```

## 6. Plot the "Applicances Energy Consumption" timeseries data using ggplot
```{r}
# Convert xts object into dataframe
df_energy = fortify(xts_energy)

# Plot the time series
ggplot(data= df_energy) +
  geom_line(aes(x = Index, y = Appliances), color = 'blue') +
  scale_x_datetime(expand = c(0,0),  labels = date_format("%m-%d", tz = 'UTC'), breaks = seq(min(index(xts_energy)), max(index(xts_energy)), by = "1 week"))+
  labs(title = "Energy Consumption Timeseries", x = "Date", y = "Energy Consumption (Wh)") +
  My_Theme
```

## 7. Plot the "Kitchen Temperature" timeseries data using ggplot
```{r}
# Plot the time series
ggplot(data= df_energy) +
  geom_line(aes(x = Index, y = T1), color = 'blue') +
  scale_x_datetime(expand = c(0,0),  labels = date_format("%m-%d", tz = 'UTC'), breaks = seq(min(index(xts_energy)), max(index(xts_energy)), by = "1 week"))+
  labs(title = "Energy Consumption Timeseries", x = "Date", y = "Energy Consumption (Wh)") +
  My_Theme
```

## 8. Decompose the "Kitchen Temperature" timeseries data into trend, seasonality, and noise parts using the stl() function which uses LOESS
```{r}
# Convert to time series object (ts) with 144 observations per day (10-minute intervals)
ts_energy_temperature = ts(as.numeric(xts_energy$T1), frequency = 144)

# Compute decomposition using the stl function
ts_energy_temperature_decomposed = stl(ts_energy_temperature, s.window = "periodic")

# Append trend to the original xts object
xts_energy =  merge(xts_energy, ts_energy_temperature_decomposed$time.series[, "trend"])
names(xts_energy)[ncol(xts_energy)] = "Trend" 

# Plot decomposition
plot(ts_energy_temperature_decomposed, main = "Decomposition of Kitchen Temperature")
```

## 9. Plot the trend and seasonality in the "Kitchen Temperature" timeseries data using ggplot
```{r}
# Convert STL output to data frame
df_energy_decomposed = data.frame(
  Index = index(xts_energy),
  Observed = as.numeric(xts_energy$T1),
  Trend = ts_energy_temperature_decomposed$time.series[, "trend"],
  Seasonal = ts_energy_temperature_decomposed$time.series[, "seasonal"],
  Remainder = ts_energy_temperature_decomposed$time.series[, "remainder"]
)

# Plot all components
ggplot(df_energy_decomposed, aes(x = Index)) +
  geom_line(aes(y = Observed, color = "Observed")) +
  geom_line(aes(y = Trend, color = "Trend"), linetype = "dashed") +
  geom_line(aes(y = Seasonal, color = "Seasonal"), linetype = "dotted")+
  scale_x_datetime(expand = c(0,0), labels = date_format("%m-%d", tz = 'UTC'), breaks = seq(min(index(xts_energy)), max(index(xts_energy)), by = "10 days"))+
   scale_color_manual(values = c("Observed" = "black", "Trend" = "blue", "Seasonal" = "red"))+
  scale_y_continuous(breaks = seq(0, 30, by = 2))+
  labs(title = "Decomposition of Kitchen Temperature (T1)",
       x = "Time", y = "Temperature (°C)") +
My_Theme
```

## 10. Extract the trend in the "Kitchen Temperature" timeseries data using LOWESS and plot it with the original data using ggplot
```{r}
model_lowess = lowess(df_energy$Index, df_energy$T1)
df_energy$Trend_Lowess = model_lowess$y

# Plot the trend extracted using LOWESS along with the original data using ggplot 
ggplot(df_energy) +
  geom_line(aes(x = Index, y = T1, color = "Observed")) +
  geom_line(aes(x = Index, y = Trend_Lowess, color = "Trend"), linetype = "dashed") +
  scale_x_datetime(expand = c(0,0), labels = date_format("%m-%d", tz = 'UTC'), breaks = seq(min(index(xts_energy)), max(index(xts_energy)), by = "10 days"))+
   scale_color_manual(values = c("Observed" = "grey", "Trend" = "blue"))+
  scale_y_continuous(breaks = seq(0, 30, by = 2))+
  labs(title = "Trend of Kitchen Temperature (T1)",
       x = "Time", y = "Temperature (°C)", color = 'Legend') +
My_Theme
```


## 11. Forecast  "Kitchen Temperature" timeseries using Holt-Winters exponential smoothing model
```{r}
# Apply Holt-Winters model
hw_model = HoltWinters(ts_energy_temperature, gamma = FALSE)

# Display model parameters
hw_model

# Forecast the next 7 days (7 * 144 = 1008 future observations)
forecast_holtwinters = forecast(hw_model, h = 1008)

# Print the forecast summary
print(forecast_holtwinters)

# Generate dynamic future timestamps
last_time = tail(index(xts_energy), 1)
future_timestamps = last_time + seq(10 * 60, 10 * 60 * length(forecast_holtwinters$mean), by = 10 * 60)
 
# Create dataframe including original and forecasted values
df_energy_forecast = data.frame(
  Index = c(index(xts_energy), future_timestamps),
  Temperature = c(df_energy$T1, forecast_holtwinters$mean),
  Type = c(rep('Original', length(index(xts_energy))), rep('Forecasted', length(future_timestamps)))
)
 
# Plot the original and forecasted values
ggplot(df_energy_forecast) +
  geom_line(aes(x = Index, y = Temperature, color = Type)) +
  scale_color_manual(values = c("Original" = "blue", "Forecasted" = "red")) +
  scale_x_datetime(expand = c(0,0), labels = date_format("%m-%d", tz = 'UTC'), breaks = seq(min(df_energy_forecast$Index), max(df_energy_forecast$Index), by = "10 days"))+
  scale_y_continuous(breaks = seq(0, 30, by = 2))+
  labs(title = "Forecast of Kitchen Temperature (T1)",
       x = "Time", y = "Temperature (°C)",  color = "Legend") +
My_Theme
```

## 12. Generate Gaussian white noise data and plot it along with its ACF and PACF
```{r}
set.seed(1)
w = rnorm(100)
y = w
Acf(y)
pacf(y)
```

## 13. Generate random walk data and plot it along with its ACF and PACF
```{r}
set.seed(1)
w = rnorm(100)
y = w
for (t in 2:100){
  y[t] = y[t-1] +w[t]
}
plot(y, type = 'l')
Acf(y)
pacf(y)
```

## 14. Difference the random walk data once and plot its ACF
```{r}
acf(diff(y))
```

## 15. Load the Adjusted Close prices for the Microsoft stock and plot using ggplot
```{r}
# Load data into xts object
getSymbols('MSFT', src = 'yahoo')
 
# Convert xts object into dataframe
df_msft = fortify(MSFT)
 
# Convert Index column to POSIXct
df_msft$Index = ymd(df_msft$Index, tz = 'UTC')
 
# Convert to xts object
xts_msft = xts(df_msft[, -1], order.by = df_msft$Index)
 
# Plot the time series
ggplot(data= df_msft) +
  geom_line(aes(x = Index, y = MSFT.Adjusted), color = 'blue')  +
   scale_x_datetime(expand = c(0,0),  labels = date_format("%Y-%m", tz = 'UTC'), breaks = seq(min(index(xts_msft)), max(index(xts_msft)), by = "1 year")) +
  labs(title = "Adjusted Close Price for MSFT Stock", x = "Date", y = "ACP") +
  My_Theme
```

## 16. Check if the Adjusted Close prices for the Microsoft stock can be modeled as a random walk
```{r}
Acf()
```

## 17. User-defined function for simulating a p-th order autoregressive process referred to as AR(p)
```{r}
generate_AR = function(n, alphas) {
  p = length(alphas)  # AR order
  w = rnorm(n)  # Generate Gaussian white noise
  y = w

  # Generate AR(p) series
  for (t in (p + 1):n) {
    y[t] <- sum(alphas * rev(y[(t - p):(t - 1)])) + w[t]
  }

  return(y)
}
```

## 18. Simulate the following autoregressive (AR) processes and check if they are stationary or not using the ACF plot
```{r}
set.seed(1)
# AR(1) with alpha_1 = 1 is simply a random walk
generate_AR(100, c(1))

# AR(1) with alpha_1 = 0.25


# AR(1) with alpha_1 = 0.6


# AR(2) with alpha_1 = 0.5 and alpha_2 = 0.5

```

## 19. Simulate an AR(2) process with alpha_1 = 0.5 and alpha_2 = 0.5, and fit an autoregressive model on the simulated data
```{r}
set.seed(1)
y = generate_AR(100, c(0.5, 0.5))
model_ar = ar(y, method = "mle")
model_ar$order
model_ar$asy.var
#model_ar$ar[1] + c(-1.96, 1.96)*sqrt(model_ar$asy.var[1,])
```

## 20. Investigate Amazon stock's Closing Price using an autoregressive (AR) model
```{r}
#getSymbols('AMZN', src = 'yahoo')
plot(AMZN$AMZN.Close)
plot(log(AMZN$AMZN.Close))
y = diff(log(AMZN$AMZN.Close))
Acf(y)
model_ar = ar(y, na.action = na.omit)
model_ar$order

```

## 22. Simulate a moving average (MA) processes and check their PACF plot
```{r}

```

## 23. Investigate Amazon stock's Closing Price using a moving average (MA) model
```{r}

```

## 24. Investigate Amazon stock's Closing Price using an AR-integrated MA (ARIMA) model
```{r}

```






