---
title: "Coding Assignment on Random Variables"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(dplyr)
# Following package is for side-by-side plots. First run install.packages('patchwork') from console
library(patchwork) 
theme_set(theme_dark())
```

#### Run the following cell to simulate the heights (in cm) of 100000 male students
```{r}
# Parameters
mu_Male = 170
sigma_Male = 8

# Bin width
delta = 2.0

nsimulations = 1e5
simulatedData_Male = rnorm(nsimulations, mean = mu_Male, sd = sigma_Male)
```

```{r}
simulatedData_Male1 = rnorm(100, mean = mu_Male, sd = sigma_Male)
simulatedData_Male1
```


#### Convert the simulated data into a data frame
```{r}
dfHeights = data.frame(simulatedData_Male)
colnames(dfHeights) = c('MaleHeight')
head(dfHeights, 3)
```

#### Calculate the following:
1. probability that a random student's height is equal to 170 cm.

```{r}
height_170cm = dfHeights$MaleHeight == 170
height_170cm
```

```{r}
height_170cm = sum(dfHeights$MaleHeight == 170)/nrow(dfHeights)
height_170cm
mean(dfHeights$MaleHeight == 170)
```

2. probability that a random student's height is in between 170 cm and 172 cm.

```{r}
height_170_172cm = sum(170 <= dfHeights$MaleHeight & dfHeights$MaleHeight <= 172)/nrow(dfHeights)
height_170_172cm
mean(170 <= dfHeights$MaleHeight & dfHeights$MaleHeight <= 172)
```

3. probability that a random student's height is in between 170 cm and 171 cm.

```{r}
height_170_171cm = sum(170 <= dfHeights$MaleHeight & dfHeights$MaleHeight <= 171)/nrow(dfHeights)
height_170_171cm
mean(170 <= dfHeights$MaleHeight & dfHeights$MaleHeight <= 171)
```

#### Plot three different types of histogram for the male students' heights and explain what each histogram means:
1. frequency histogram
2. relative frequency histogram
3. density histogram 

```{r}
# Plot the frequency histogram for male student heights
p1 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(count)),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  labs(x = 'Height (cm)', y = 'Frequency')
p1
```

```{r}
# Plot the relative frequency histogram for male student heights
p2 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(count)/sum(after_stat(count))),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  labs(x = 'Height (cm)', y = 'Relative Frequency')
p2
```

```{r}
# Plot the density histogram for male student heights
p3 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(density)),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  labs(x = 'Height (cm)', y = 'Density')
p3
```

```{r}
# Plot the density histogram along with density curve for male student heights
p4 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(density)),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  stat_function(fun = dnorm, args = list(mean = mean(dfHeights$MaleHeight), sd = sd(dfHeights$MaleHeight))) +
  labs(x = 'Height (cm)', y = 'Density')
p4
```


```{r}
# Plot the frequency histogram for male student heights
p1 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(count)),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  labs(x = 'Height (cm)', y = 'Frequency')

# Plot the relative frequency histogram for male student heights
p2 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(count)/sum(after_stat(count))),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  labs(x = 'Height (cm)', y = 'Relative Frequency')

# Plot the density histogram for male student heights
p3 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(density)),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  labs(x = 'Height (cm)', y = 'Density')

# Plot the density histogram along with density curve for male student heights
p4 = ggplot(data = dfHeights) +
  geom_histogram(aes(x = MaleHeight, y = after_stat(density)),
  breaks = seq(mu_Male-4*sigma_Male, mu_Male+4*sigma_Male, by = delta),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  stat_function(fun = dnorm, args = list(mean = mean(dfHeights$MaleHeight), sd = sd(dfHeights$MaleHeight))) +
  labs(x = 'Height (cm)', y = 'Density')


# Plot the histograms side-by-side
p1 + p2 + p3 + p4 + plot_layout(ncol = 2)
```

```{r}
height_170_172cm = sum(156 <= dfHeights$MaleHeight & dfHeights$MaleHeight <= 160.5)/nrow(dfHeights)
height_170_172cm
mean(156 <= dfHeights$MaleHeight & dfHeights$MaleHeight <= 160.5)
```

