---
title: "Coding Assignment on dplyr and ggplot"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(dplyr)
```

#### Load the bodyfat dataset
```{r}
# Load the bodyfat dataset
file = 'Data/bodyfat.csv'
df = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
str(df)
head(df, n = 2)
```

#### Create lists of categorical and continuous features
```{r}
# Create lists of categorical and continuous features
features = colnames(df)
categorical_features = c('Gender')
continuous_features = features[!(colnames(df) %in% categorical_features)]
```

#### Convert categorical columns to correct type
```{r}
# Convert categorical columns to correct type
df[categorical_features] = lapply(df[categorical_features], as.factor)
str(df)
```

#### Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles
```{r}
# Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles
df %>% summarise(across(-Gender, mean))
mean(df$Weight) 
df$Weight - mean(df$Weight) # Deviations vector
(df$Weight - mean(df$Weight))^2 # Squared - Deviations vector
mean(df$Weight - mean(df$Weight)) # Sample varinace
var(df$Weight) # Variance in-bult
sqrt(mean((df$Weight - mean(df$Weight))^2))#standard deviation
sd(df$Weight) # Standard Deviation in-bult
(df$Weight - mean(df$Weight))/sd(df$Weight)#standardized weight value with no units(Z-score)
scale(df$Weight)#standardized weight value with no units(Z-score) in-bult

```

#### Plot the distribution of males and females in the dataset
```{r}
# Plot the distribution of males and females in the dataset
p = ggplot(data = df) +
  geom_bar(aes(x = Gender)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Gender',
       y = 'Frequency',
       title = 'Distribution of Genders')
p

```


#### Plot a component plot of weight values
```{r}
# Plot a component plot of weight values
p = ggplot(data = df) +
  geom_point(aes(x = 1:nrow(df),y = Weight)) + geom_hline(yintercept = mean(df$Weight), color = 'red', linetype = 'dashed')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Weights')
p
```

#### Plot a component plot of Bodyfat values
```{r}
# Plot a component plot of weight values
p = ggplot(data = df) +
  geom_point(aes(x = 1:nrow(df),y = Bodyfat)) + geom_hline(yintercept = mean(df$Bodyfat), color = 'red', linetype = 'dashed')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Bodyfat',
       title = 'Component Plot of Bodyfat')
p
```

#### Plot a component plot of Mean - Centered Bodyfat values
```{r}
# Plot a component plot of weight values
MeanCentered = df$Bodyfat - mean(df$Bodyfat)
p = ggplot(data = df) +
  geom_point(aes(x = 1:nrow(df),y = Bodyfat - mean(Bodyfat))) + geom_hline(yintercept = mean(df$Bodyfat), color = 'red', linetype = 'dashed')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Mean - Centered Bodyfat')
p
```

#### Plot a histogram of weight values
```{r}
#### Plot a histogram of weight values
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
geom_histogram(aes(x=Weight,y=after_stat(count)), breaks=seq(mean_weight-3*sd_weight,mean_weight+3*sd_weight)
       ,binwidth=1,color='black',fill='steelblue',alpha=0.4)+
geom_vline(xintercept = mean(df$Weight),color='red',linetype="dashed") +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights')
p
```

#### Plot a histogram of height values
```{r}
# Plot a histogram of height values
mean_height = mean(df$Height)
sd_height = sd(df$Height)
p = ggplot(data = df) +
geom_histogram(aes(x=Height,y=after_stat(count)), breaks=seq(mean_height-3*sd_height,mean_height+3*sd_height)
       ,binwidth=1,color='black',fill='steelblue',alpha=0.4)+
geom_vline(xintercept = mean(df$Height),color='red',linetype="dashed") +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Height',
       y = 'Frequency',
       title = 'Frequency Histogram of Height')
p
```

#### Plot a histogram of bodyfat values
```{r}
# Plot a histogram of bodyfat values
mean_Bodyfat = mean(df$Bodyfat)
sd_Bodyfat = sd(df$Bodyfat)
p = ggplot(data = df) +
geom_histogram(aes(x=Bodyfat,y=after_stat(count)), breaks=seq(mean_Bodyfat-3*sd_Bodyfat,mean_Bodyfat+3*sd_Bodyfat)
       ,binwidth=1,color='black',fill='steelblue',alpha=0.4)+
geom_vline(xintercept = mean(df$Bodyfat),color='red',linetype="dashed") +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Bodyfat',
       y = 'Frequency',
       title = 'Frequency Histogram of Bodyfat')
p
```

#### Plot a histogram of weights for both genders
```{r}
# Plot a histogram of weights for both genders
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
  geom_histogram(aes(x = Weight, y= after_stat(count),color=Gender),
                 breaks = seq(mean_weight-3*sd_weight,mean_weight +3*sd_weight,by = 2.0),alpha = 0.9)+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights')
p
```


#### Plot a boxplot of bodyfat values
```{r}
# Plot a boxplot of bodyfat values
p = ggplot(data = df) + geom_boxplot(aes( y = Bodyfat)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = '',
       y = 'Bodyfat',
       title = 'Frequency Histogram of Weights')
p
```

#### Plot a boxplot of bodyfat values for both genders
```{r}
# Plot a boxplot of bodyfat values for both genders
p = ggplot(data = df) + geom_boxplot(aes(x = Gender, y = Bodyfat)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Bodyfat',
       y = 'Gender',
       title = 'Frequency Histogram of Weights')
p
```

#### Introduce a new categorical column in the dataframe called 'Weightlevel' with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)
```{r}
# Introduce a new categorical column in the dataframe called weight level with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)
q1 = quantile(df$Weight, 0.25)
q2 = quantile(df$Weight, 0.75)
df = df %>% mutate(Weightlevel = case_when(Weight < q1 ~ "Low" , Weight > q2 ~ "High" , Weight > q1 & Weight < q2 ~ "Normal" ))
df
```

#### Plot a boxplot of bodyfat values for both genders across the three Weightlevel groups
```{r}
# Plot a boxplot of bodyfat values for both genders both genders across the three Weightlevel groups
p = ggplot(data = df) + geom_boxplot(aes(x = Gender, y = Bodyfat, fill = Weightlevel)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Gender',
       y = 'Bodyfat',
       title = 'Frequency Histogram of Weights')
p
```

#### The 68-95-99 rule
```{r}
# The 68-95-99 rule

```


