---
title: "Intro to R"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Load the food texture dataset
```{r}
#file = 'http://openmv.net/file/food-texture.csv'
file = 'Data/food-texture.csv'
foodData = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
head(foodData, 2)
str(foodData)
```

#### Modify Crispy column to reflect high (0) or low (1) crispiness
```{r}
foodData = foodData %>% mutate(Crispylevel = ifelse(Crispy > 11, 'High', 'Low'))
head(foodData)
```

##### Change Crispy and Crispylevel columns to factor (categorical) type
```{r}
# Continuous features -> Oil, Density, Hardness, Fracture
# Categorical features -> Crispy (8 levels 8 through 15) and Crispylevel (2 levels 0 and 1)
# Crispy is a categorical feature with an order
str(foodData)
categorical_cols = c('Crispy', 'Crispylevel')
foodData[categorical_cols] = lapply(foodData[categorical_cols], as.factor)
str(foodData)
```

#### Visualize the OilPercentage feature using a histogram
```{r}
histogram = ggplot(data = foodData) + geom_histogram(aes(x = Oil), binwidth = 0.1, color = 'blue')
histogram
```

#### Scatter plot between OilPercentage and Density
```{r}
scatter = ggplot(data = foodData) + geom_point(aes(y = Oil, x = Density), color = 'red') + labs(x = 'Density', y = 'OilPercentage', title =  'OilPercentage and Density')
scatter
```

#### Scatter plot between OilPercentage and Density color coded with Crispylevel
```{r}
scatter1 = ggplot(data = foodData) + geom_point(aes(y = Oil, x = Density, color = Crispylevel)) + labs(x = 'Density', y = 'OilPercentage', title =  'OilPercentage and Density')
scatter1
```

#### Scatter plot between Density and Hardness
```{r}
scatter2 = ggplot(data = foodData) + geom_point(aes(y = Density, x = Hardness), color = 'black') + labs(x = 'Density', y = 'Hardness')
scatter2
```

#### Scatter plot between Density and Fracture
```{r}
scatter3 = ggplot(data = foodData) + geom_point(aes(y = Density, x = Fracture), color = 'steelblue') + labs(x = 'Density', y = 'Fracture')
scatter3
```

#### Smooth line plot using ggplot
```{r}
lineplot = ggplot(data = foodData) + geom_smooth(aes(y = Oil, x = Density, color = 'red'))
lineplot
```

#### Barplot for Crispy
```{r}
Barplot = ggplot(data = foodData) + geom_bar(aes(x = Density), color = 'green')
Barplot
```

#### In-built functions for dataframes
```{r}
# Mean oil percentage across all samples

# Mean-centering of OilPercentage

# Sum of the squared deviation from the mean

# Average of the squared deviation from the mean

# Variance of OilPercentage

# Standard deviation of OilPercentage
```


#### Box plot using ggplot with color coding
```{r}
Box = ggplot(data = foodData) + geom_boxplot(aes(y = Oil),color = 'red')
Box
```


```{r}
Box1 = ggplot(data = foodData) + geom_boxplot(aes(x = Crispylevel, y = Oil),color = 'blue')
Box1
```

```{r}
foodData %>% select(Oil)
Oil_Median = median(foodData$Oil) # 50th percentile or 0.5th quantile
mean(foodData$Oil)
```

```{r}
foodData %>% filter(Oil <= Oil_Median) %>% select(Oil)
nrow(foodData %>% filter(Oil <= Oil_Median) %>% select(Oil))/nrow(foodData)
mean(foodData$Oil <= Oil_Median)
```













